---
title: "Assignment 1st"
author: " Shubham Bhargava:sb2509@scarletmail.rutgers.edu"
date: "09/24/2023"
output: html_document
---

# Typical Industries Employee Survey

```{r, message=FALSE}
# Reading the excel file
library(readxl)
Typical_Employee_Survey_Data <- read_excel("C:/Users/HP/Downloads/Typical_Employee_Survey_Data.xlsx")
View(Typical_Employee_Survey_Data)

```


#### 1. Creating header Row
```{r}
names(Typical_Employee_Survey_Data) <- c('Age','Gender','Job_Satisfaction','Job_Preference','years_with_current_employer','Promotion_Likelihood','Decision_making_Involvement','Budget_Participation','Job_Pride','Job_Loyalty','Workplace_relation')
names(Typical_Employee_Survey_Data)


```


#### 2.Print out the 5-point summary and boxplot for each column

```{r}
summary(Typical_Employee_Survey_Data)
boxplot(Typical_Employee_Survey_Data$Age, main = 'Age of people', ylab = 'Range of Age', col = 'Red')
boxplot(Typical_Employee_Survey_Data$years_with_current_employer, main = 'Years with current Employer', ylab = 'Range of years', col = 'Red')
#boxplot for others column as type is numeric, before factoring
boxplot(Typical_Employee_Survey_Data$Gender, main = 'Gender of people', ylab = 'Range of Gender from 1 to 2', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Job_Satisfaction, main = 'Job Satisfaction', ylab = 'Range of Satisfaction', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Job_Preference, main = 'Imp Job Characteristics', ylab = 'Range of Preference', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Promotion_Likelihood, main = 'Likelihood of getting promoted', ylab = 'Range of Promotion Likelihood', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Decision_making_Involvement, main = 'Involvement in making decision', ylab = 'Range of decision involvement', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Budget_Participation, main = 'Participation in Budgetary decision', ylab = 'Range of Budgetary Decision', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Job_Pride, main = 'Pride level working for the Organization', ylab = 'Range of pride level', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Job_Loyalty, main = 'Loyalty level for the Organization', ylab = 'Range of loyalty level', col = 'Red')
boxplot(Typical_Employee_Survey_Data$Workplace_relation, main = 'Level of workplace relation', ylab = 'Range of relation level', col = 'Red')




```


#### 3.Convert the columns that need to be converted to Factors

```{r}
Typical_Employee_Survey_Data$Gender <- factor(Typical_Employee_Survey_Data$Gender, levels = c(1, 2), labels = c("Male", "Female"))
Typical_Employee_Survey_Data$Job_Satisfaction <- as.factor(Typical_Employee_Survey_Data$Job_Satisfaction)
Typical_Employee_Survey_Data$Job_Preference <- as.factor(Typical_Employee_Survey_Data$Job_Preference)
Typical_Employee_Survey_Data$Promotion_Likelihood <- as.factor(Typical_Employee_Survey_Data$Promotion_Likelihood)
Typical_Employee_Survey_Data$Decision_making_Involvement <- as.factor(Typical_Employee_Survey_Data$Decision_making_Involvement)
Typical_Employee_Survey_Data$Budget_Participation <- as.factor(Typical_Employee_Survey_Data$Budget_Participation)
Typical_Employee_Survey_Data$Job_Pride <- as.factor(Typical_Employee_Survey_Data$Job_Pride)
Typical_Employee_Survey_Data$Job_Loyalty <- as.factor(Typical_Employee_Survey_Data$Job_Loyalty)
Typical_Employee_Survey_Data$Workplace_relation <- as.factor(Typical_Employee_Survey_Data$Workplace_relation)
str(Typical_Employee_Survey_Data)
summary(Typical_Employee_Survey_Data)

```


#### 4.Plot the right visual for each column

```{r}
hist(Typical_Employee_Survey_Data$Age,main = 'Employee Age', ylab='Number of Employees', xlab='Frequncy of Age',col='Yellow',border='orange', xlim = c(20,70))

#* Pie chart for Gender distribution
pie_data <- table(Typical_Employee_Survey_Data$Gender)
total_count <- sum(pie_data)
labels <- paste(c("Male","Female"), ": ", pie_data,"(",round(pie_data/total_count*100,1),"%)")
pie(pie_data,main = "Gender Distribution",label= labels,col = c("black", "purple"))

#*barplot for Job satisfaction
barplotJS <- table(Typical_Employee_Survey_Data$Job_Satisfaction)
barplot(barplotJS , main='Job Satisfaction Plot', names.arg= c("Very Satisfied","Moderately Satisfied","Dissatisfied","Very Dissatisfied"),ylab = 'count of Employees', xlab='Satisfaction level', col = 'lightBlue', ylim=c(0,70))

#* Job Preference plot
pie_dataJP <- table(Typical_Employee_Survey_Data$Job_Preference)
total_countJP <- sum(pie_dataJP)
labels <- paste(c("High Income","No Danger of Being Fired","Flexible Hours","Opportunities for Advancement","Enjoying the Work"), ": ", pie_dataJP,"(",round(pie_dataJP/total_countJP*100,1),"%)")
pie(pie_dataJP,main = "Imp Job Characteristic",label= labels,col = c("black", "purple",'grey','blue','orange'))


#* Working years with current employer
hist(Typical_Employee_Survey_Data$years_with_current_employer,main = 'Years with current Employer', ylab = 'Number of Employees', col = 'lightBlue', xlab='Range of Experience', xlim=c(0,60),ylim=c(0,70))

#*how likely to get promoted
barplotLP <- table(Typical_Employee_Survey_Data$Promotion_Likelihood)
barplot(barplotLP , main='Likelihood of getting promoted', names.arg= c("Very Likely","Likely","Not Sure","Unlikely","Very Unlikely"),ylab = 'Number of Employees', xlab='likelihood level', col = 'lightBlue')

#*part in making decisions
pie_dataDMI <- table(Typical_Employee_Survey_Data$Decision_making_Involvement)
total_countDMI <- sum(pie_dataDMI)
labels <- paste(c("Always","Much of the Time","Sometimes","Never"), ": ", pie_dataDMI,"(",round(pie_dataDMI/total_countDMI*100,1),"%)")
pie(pie_dataDMI,main = "Involvement in making decision",label= labels,col = c("orange", "blue",'purple','black','white'))

#*participate in budgetary decisions
barplotBD <- table(Typical_Employee_Survey_Data$Budget_Participation)
barplot(barplotBD , main='Participation in Budgetary decision', names.arg= c("Yes","No"),ylab = 'Number of Employees', xlab='Budgetary decision', col = 'lightBlue', ylim=c(0,70))

#*pride level working for this organization
pie_dataJbP <- table(Typical_Employee_Survey_Data$Job_Pride)
total_countJbP <- sum(pie_dataJbP)
labels <- paste(c("Very Proud","Somewhat Proud","Indifferent","Not At All Proud"), ": ", pie_dataJbP,"(",round(pie_dataJbP/total_countJbP*100,1),"%)")
pie(pie_dataJbP,main = "Pride level working for the Organization",label= labels,col = c("black", "purple",'yellow','orange'))

#*Loyalty with this organization
barplotJL <- table(Typical_Employee_Survey_Data$Job_Loyalty)
barplot(barplotJL , main='Loyalty level for the Organization', names.arg= c("Very Likely","Likely","Not Sure","Unlikely","Very Unlikely"),ylab = 'Number of Employees', xlab='Loyalty Level', ylim=c(0,70))

barplotWR <- table(Typical_Employee_Survey_Data$Workplace_relation)
barplot(barplotWR , main='Level of workplace relation', names.arg= c("Very Good","Good","So So","Bad"),ylab = 'Number of Employees', xlab='Work relation Level', ylim=c(0,60))

```

#### 5.Plot scatter plots to show few relationships

*As numerical columns available are year of employment and Employees Age Scatter plot is drawn
```{r}
# Highest density is present between employee age 30 to 40 yrs
plot(Typical_Employee_Survey_Data$Age,Typical_Employee_Survey_Data$years_with_current_employer, main="Employee Age VS. year of employment", xlab="Employee Age", ylab="year of employment",col='blue',pch=19)
```

*From the given plot we can infer that majority of workforce participating in survey, 
comprises of people within the age 20-40 with overall experience of less that 10 years within company.

```{r}
#Below are additional relations plot:
#year of employment and promotion likelihood
plot(Typical_Employee_Survey_Data$Promotion_Likelihood,Typical_Employee_Survey_Data$years_with_current_employer, main="year of employment VS. promotion likelihood", xlab="Promotion likelihood Rating 1 being Very Likely", ylab="year of employment",col='black')

#year of employment Vs. Job Pride
plot(Typical_Employee_Survey_Data$Job_Pride,Typical_Employee_Survey_Data$years_with_current_employer, main="year of employment VS. Job Pride", xlab="Job Pride Rating 1 being Very Proud", ylab="year of employment",col='grey')

#year of employment and Job Loyalty
plot(Typical_Employee_Survey_Data$Job_Loyalty,Typical_Employee_Survey_Data$years_with_current_employer, main="year of employment Vs. Job Loyalty", xlab="Job Loyalty 1 being Very Likely", ylab="Years of Employment",col='purple')










```
